Continuous developmental change can explain discontinuities in word learning

### 2019-07-22 journal/ms.Rmd
journal v1

 "Cognitive development is often characterized in term of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using simpler testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model describing how development may proceed in a continuous fashion across the lifespan.  The model accounts for previously documented facts and provides new predictions.  We collected data from preschool children and adults, and we showed that the model can explain various patterns of learning both within the same age and across development. The findings suggest that major aspects of cognitive development that are typically thought of as discontinuities, may emerge from simpler, continuous mechanisms."

@@

```{r load_packages, include = FALSE}
library(papaja)
library(png)
library(grid)
library(ggplot2)
library(ggthemes)
library(xtable)
options(xtable.comment = FALSE)
library(purrr)
library(readr)
library(ggplot2)
library(langcog)
library(boot)
#library(lazyeval)
library(dplyr)
library(tidyr)
library(wordbankr)
library(directlabels)
#library(scales)
library(stringr)
library(lmtest)
library(rwebppl)
library(jsonlite)
library(nlme)
library(feather)
library(broom)
library(HDInterval)
library(BBmisc)
library(stargazer)
library(lme4)
```

@@

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Indeed, though infants around 14-month old can distinguish similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects.

<!--Follow-up studies has focused on proposing possible explanations for this observed gap between speech perception and word learning [e.g., @fennell2010; @stager1997; @hofer2017; @rost2009]. -->

By around 17 m.o, children succeed in the same task [@werker2002]. How does development proceed? The answer to this question depends on the nature of the chidlren's word encoding in the task. Early accounts assumed that children encode words in a rather binary way: they either fail or succeed in attending to the relevant phonetic details (simultaneously with the meanings). In the orginal paper, @stager1997 noted that "Infants of 14 months fail to detect the same phonetic detail that they can easily detect in a simple syllable discrimination task". They argued that the high cogntive demands of the task make children experience a limitation in the amount of phonetic details they can access. In a similar vein, Pater et al. (2004) argued that "when a contrast is first acquired, it is not stable and can be lost under processing demands". This binary account suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children succeed.

Subsequent findings have investigated the implications of this hypothesis in both infants and adults. First, if the demands of the task totally empede the encoding of subtle phonetic contrasts, then infants should fail in learning the mappings between the sounds and the objects, regardless of how this learning is probed. However, 14-month-olds succeed in this task under the same *learning* conditions as in @stager1997, but when an easier *testing* method is used [@yoshida2009]. Second, if the minsmatch between sound distrimination and word learning is only related to limited cognitive resources in infancy, then adults --- who have greater attentianal and working memory capacity --- should always succeed in mapping pairs of sounds they can discriminate to different objects.  Nonetheless, even adults show patterns learning that mirror those shown by 14-month-olds [@white2013; @pajak2016].

This new set of evidence points towards another scenario, where the representations are encoded in a probabilistic (rather than binary) way, and where development is continuous, rather than stage-like [see also @swingley2007]. On this account, correct representations are learned early in development, but these representations are encoded with higher uncertainty in younger children, leading to *apparent* failure in relatively demanding tasks. Development is a continuous process whereby the initial noisy representations become more precise. Crucially, in a probabilsitic account, more precise representation do not mean they perfect, thus accounting for the fact that even adults show low accuracy learning when the sounds are subtle, e.g., non-native [@pajak2016].

We provide an intuitive illustration of how such an account explains patterns of learning and development in Figure \@ref(fig:illus). We observe low accuracy in word learning when the perceptual distance between the labels is small relative to the uncertainty with which these labels are encoded. For example, in Stager and Werker's original experiment, children are supposed to associate label 1 ("bih") and label 2 ("dih") with object 1 and object 2, respectively.  Though infants could learn that the label "bih" is a better match to object 1 than "dih", they could still judge the sound "dih" as a plausible instance of the label "bih", thanks to the relatively large uncertainty of the encoding, and this confusion leads to "failure" in the recognition task. According to this account, accuracy in word learning improves if we increase either the perceptual distinctiveness of the stimuli (e.g., through using different-sounding labels), or the precision of the encoding itself (e.g., across development).

Building on this intuition, the current work proposes a probabilistic model, which we use to both account for previous experimental findings, and to make new predictions that have not been tested before. Using new data collected from both preschool children and adults, we show that the model can explain various patterns of learning both within the same age and across development.

@@

The learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound. To this end, they must compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$

@@

\begin{equation}
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\end{equation}

@@

We use the model to predict performance in the word learning task introduced by @stager1997, with a two-alternative forced choice as in @yoshida2009. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \@ref(fig:task).

We used Equation 1 and the probability distributions defined above to obtain the exact analytical expression for the probability of accurate responses $p(o_T | s)$ (target object $o_T$ given a sound $s$) in the simple case of two-alternative forced choice in the testing phase of our experimental task:

\begin{equation}
P(o_T|s)= \frac{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}}{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}+ e^{-\Delta s^2 /2\sigma_L^2} + e^{-\Delta o^2 /2\sigma_C^2 }}
\end{equation}

Figure \ref{fig:simulation} show simulations of the predicted accuracy (Expression 2) as a function of the distinctiveness parameters ($\Delta s$ and $\Delta o$) and the precision parameters, i.e., the variances of the distributions $p(s| L)$ and $p(o | C)$. To understand the qualitative behavior of the model, we assumed for simplicity that the precision parameter has similar values in both distributions, i.e., $\sigma =\sigma_C \approx \sigma_L$ (but we will allow those parameters to vary independently in the rest of the paper).

@@

The simulations explain some previously documented facts, and make new predictions:

@@

2) For fixed values of $\Delta s$ and $\Delta o$, accuracy increases when the representational uncertainty (characterized with $\sigma$) decreases. This fact may explain development, i.e., younger children have noisier representations [see @swingley2007; @yoshida2009], which leads to lower word recognition accuracy, especially for similar-sounding words.

3) For fixed values of $\Delta s$ and $\sigma$, accuracy increases with the visual distance between the semantic referents $\Delta o$. This is a new prediction that our model makes. Previous work studied the effect of several bottom-up and top-down properties in disambiguating similar sounding words [e.g., @fennell2010; @rost2009; @thiessen2007], but to our knowledge no previous study in the literature tested the effect of the visual distance between the semantic referents.

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \@ref(fig:task)). More precisely, we explored the predictions related to both distinctiveness and precision. Sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameters, $\sigma_C$ and $\sigma_L$. The experiment, sample size, exclusion criteria and the model's main predictions were pre-registered.

@@

d_child <- read_delim("../Analysis/kids_all.txt", delim = " ") %>%
  filter(type == "Test") %>%
  filter(age != 6) %>%
  filter(code != 'test',
         !is.na(code)) %>%
  filter(grepl('soph', code)) %>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

@@

### Participants
We planned to recruit a sample of N=60 children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. Here we report data from n=`r N_good_ch` children. An additional n=`r N_all_ch-N_good_ch` children participated but were removed from analyses because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also planned to recruit a sample of N=60 adults on Amazon Mechanical Turk. Data from n=`r N_all_ad-N_good_ad` participants were excluded due to low scores on the catch trials (n=`r N_all_ad-N_good_ad-d_adult_lang` ) or because they were familiar with the non-English sound stimuli we used in the adult experiment (n=`r d_adult_lang`), yielding a final sample of n=`r N_good_ad`.

### Stimuli and similarity rating
The sound stimuli were generated using the MBROLA Speech Synthesizer [@dutoit1996]. We generated three kinds of nonsense word pairs which varied in their degree of similarity to English speakers: 1) "different": "lif"/"neem" and "zem"/"doof", 2) "intermediate": "aka"/"ama" and "ada"/"aba", and 3) "similar" non-English minimal pairs: "ada"/"a\textipa{d\super h}a" (in hindi) and "a\textipa{Q}a"/"a\textipa{\textcrh}a" (in arabic).

As for the objects, we used the Dynamic Stimuli javascript library\footnote{https://github.com/erindb/stimuli} which allowed us to generate objects in four different categories: "tree", "bird", "bug", and "fish". These categories are supposed to be naturally occurring kinds that might be seen on an alien planet. In each category, we generated "different", "intermediate" and "similar" pairs by manipulating a continuous property controlling features of the category's shape (e.g, body stretch or head fatness).

In a separate survey, $N=20$ participants recruited on Amazon Mechanical Turk evaluated the similarity of each sound and object pair on a 7-point scale. We scaled responses within the range [0,1]. Data are shown in Figure \@ref(fig:stim), for each stimulus group. These data will be used in the models as the perceptual distance of sound pairs ($\Delta s$) and object pairs ($\Delta o$).

@@

#Import data from saved
data_human_model <- feather::read_feather("../saved_data/data_human_model.feather") #%>%
  #mutate(Model = ifelse(key == "prediction", "Single Variance", "Double Variance"))

@@

plot_data
```
###Design

Each age group saw only two of the three levels of similarity described in the previous sub-section: "different" vs. "intermediate" for preschoolers, and "intermediate" vs. "similar" for adults. We made this choice in light of pilot studies showing that adults were at ceiling with "different" sounds/objects, and children were at chance with the "similar" sounds/objects. That said, this difference in the level of similarity is accounted for in the model through using the appropriate perceptual distance used in each age group (Figure \@ref(fig:stim)).

@@

To maximize our ability to measure subtle stimulus effects, the experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the 4 conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones, but was used only to select participants who were able to follow the instructions and show minimal learning.

@@

```{r allData, fig.env = "figure", fig.pos = "h", fig.align='center', fig.cap = "Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\\% confidence intervals."}

@@

Experimental results are shown in Figure \@ref(fig:allData) (solid lines). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings. For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that performance improves with age. In addition, we found two-way interactions between sound distance and age ($\beta =$ `r s_age_coef`, $p$ `r p_val(s_age_p)`) and between object distance and age ($\beta =$ `r c_age_coef`, $p$ `r p_val(c_age_p)`).

@@

R2_model1 <-  round(cor(data_correlations$human, data_correlations$prediction)^2, 2)
R2_model2 <-  round(cor(data_correlations$human, data_correlations$prediction_2)^2, 2)

@@

We next fit our model (using Equation 2) to the participants' responses in each age group.  The values of $\Delta s$ and $\Delta o$ were set based on data from the similarity judgment task (Figure \@ref(fig:stim)). Thus, the model has two degrees of freedom for each group, i.e., $\sigma_C$ and $\sigma_L$.  Figure \@ref(fig:allData) (dashed lines) shows the predictions. The model captures the qualitative patterns in both age groups: starting from a low accuracy recognition when both the sound and object distances are small, the model correctly predicts an increase in accuracy when either the sound distance or the object distance increases. Further, accuracy is correctly predicted to be maximal when both the sound and object distances are high.

The values of the parameters were as follows. Children had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)`]\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.}, and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)`]. Adults had a label-specific uncertainty of $\sigma_S =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)`], and a concept-specific uncertainty of $\sigma_C =$ `r round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)` [`r round(mod2_val$lower[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`, `r round(mod2_val$upper[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)`].  As predicted, the uncertainty parameters were larger for children than they were for adults, showing that the probabilistic representations becomes more refined (that is, $\sigma$ becomes smaller) across development. The developmental effect was more important for the label-specific uncertainty.

The models explained the majority of the variance in the participants' mean responses ($R^2=$ `r R2_model2`, for the combined adult and children's data). To investigate whether the model's predictive power was due to overfitting, we fit a simplified version with only one degree of freedom (i.e., a single variance common to both sounds and objects). This single-variance model explained as much variance in the mean responses ($R^2=$ `r R2_model1`). It also captured the main qualitative patterns (graph not shown), suggesting that the explanatory power of the model is largely due to its structure, rather than its degrees of freedom.

>>>>

 "Cognitive development is often characterized in term of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using simpler testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model describing how development may proceed in a continuous fashion across the lifespan.  The model accounts for previously documented facts and provides new predictions.  We collected data from preschool children and adults, and we showed that the model can explain various patterns of learning both within the same age and across development. The findings suggest that major aspects of cognitive development that are typically thought of as discontinuities may emerge from simpler, continuous mechanisms."

@@

```{r load_packages, include = FALSE}
library(papaja)
library(png)
library(grid)
library(ggplot2)
library(ggthemes)
library(xtable)
options(xtable.comment = FALSE)
library(purrr)
library(readr)
library(ggplot2)
library(langcog)
library(boot)
#library(lazyeval)
library(dplyr)
library(tidyr)
library(wordbankr)
library(directlabels)
#library(scales)
library(stringr)
library(lmtest)
library(rwebppl)
library(jsonlite)
library(nlme)
library(feather)
library(broom)
library(HDInterval)
library(BBmisc)
library(stargazer)
library(lme4)
library(kableExtra)
```

@@

We use a case study from word learning literature. @stager1997 first showed that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. The authors measured word learning using an audio-visual habituation Switch task. First, infants are familiarized with two word-object pairings (e.g., Word A with Object A and Word B with Object B). Second, they are tested using two types of trials. The control 'same' trial consists of a correct pairing  (e.g., Word A with Object A) and the 'switch' trial consists of a wrong pairing (e.g., Word A with Object B). If babies have correctly learned the association during the familiarization, they are supposed to be surprised by the 'switch' trial and not by the 'same' trial. The former should thus result in a greater looking time compared to the latter [@werker1998].

Though infants around 14-month old can distinguish perceptually similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects in the switch task. This failure has initially been taken as evidence that 14-month olds do not encode subtle sounds during meaning learning [@stager1997; @pater2004]. This interpretation suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children, around 17 months [@werker2002], typically do.

<!-- In fact, @stager1997 noted that "Infants of 14 months fail to detect the same phonetic detail that they can easily detect in a simple syllable discrimination task". They argued that the high cogntive demands of the task make children experience a limitation in the amount of phonetic details they can access. In a similar vein, Pater et al. (2004) argued that "when a contrast is first acquired, it is not stable and can be lost under processing demands"-->

<!--Follow-up studies has focused mainly on proposing possible explanations for the observed gap between speech perception and word learning [e.g., @fennell2010; @stager1997; @hofer2017; @rost2009]. -->

The initial discontinuous interpretation has been challenged by subsequent work. For instance, @yoshida2009 investigated whether failure in the switch task reflects a lack of sound encoding during *familiarization*, or whether it is only due to the demands of the *testing* method which does not allow learning below a certain threshold to be detected. They used the same familiarization procedure as @stager1997, but instead of comparing the looking times in 'same' and 'switch' trials, they tested infants using a two-alternative choice task comparing fixations to target and distractor objects [@golinkoff1987; @fernald2006]. Using this simpler and finer-grained task, the researcher found evidence for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from adult studies. If the mismatch between sound discrimination and word learning is only a stage in early infancy, then this mismatch should disappear by adulthood. Nonetheless, even adults show patterns of learning that mirror those shown by 14-month-olds when the sound contrasts more challenging [@white2013; @pajak2016].

Some researchers [e.g., @swingley2007; @yoshida2009; @pajak2016] proposed that the phonological form may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding the phonological form of words in a probabilistic fashion. According to this view, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy representations.

In a probabilistic account, a word can be represented as a probability distribution over sound instances organized in a similarity space. The probability is highest at the most typical sound instance. It decreases as the instance becomes less typical. The precision of the representation can be characterized by how much it tolerates slightly atypical pronunciations. This tolerance is captured formally by the variance of the probability distribution: larger variance indicates higher tolerance and lower precision, whereas smaller variance indicates lower tolerance and higher precision.

This framework can explain several findings. In Stager and Werker's original experiment, children are supposed to associate one label "bih" with object 1 and a second label "dih" with object 2.  Though infants could learn that the label "bih" is a better match to object 1 than "dih", they could still judge the sound "dih" as a plausible instance of the label "bih", thanks to the relatively large variance of the early encoding, and this confusion leads to "failure" in the recognition task (Figure \@ref(fig:illus), top). Though learning is small and is easily disrupted by the Switch task, it can still be detected when less demanding methods are used [@yoshida2009].

The learning accuracy increases (and is detected even by demanding tests such as the Switch) for more distinct word-forms (e.g., "lif" vs. "neem") where the perceptual distance is large relative to the variance (Figure \@ref(fig:illus), left).  Distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

Finally, development can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic representations (Figure \@ref(fig:illus), right). Such an increase in precision renders minimally different sounds less confusing. Importantly, a more precise representation still has a non-zero variance --- Learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrate in adults studies [@white2013; @pajak2016].

## This study

The probabilistic account has been put forward to explain patterns of learning and development at the qualitative level. However, it is crucial to have a precise computational instantiation of this account which would allow us to *quantify* its explanatory power. We could find one previous work that attempted to provide such a computational instantiation  (Hofer and Levy, 2017). However, this previous work was designed with the goal of reproducing the results of a specific study [@pajak2016] which focused on explaining the mismatch between speech perception and word learning in adults rather than on exploring the mechanism of development.

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before (i.e., the effect of the referents' similarity on word learning). Crucially, we explored the extent to which development can be understood as a continuous refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we explain how it allows us to characterize behavior in a word learning task which resembles the one used in @stager1997 and @yoshida2009. Then we explore the predictions of the model through simulating its behavior across different parameter settings. Next we quantify the extent to which the model's predictions account for human data we collected from both preschool children and adults. Finally, we discuss the results in the lights of existing accounts of word development.

@@

The learner hears a sound $s$ and has to decide which object $o$ provides an optimal match to this sound (see Figure \@ref(fig:task)). To this end, they must compute the probability $P(o|s)$ for all possible objects. This probability can be computed by summing over all possible concepts and labels:
$$P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) $$

@@

$$
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
$$

@@

We use the model to predict word learning in a task similar to the one introduced by @stager1997. We used a modified version of the task where the testing method consists in a two-alternative forced choice [@yoshida2009]. In this task, participants are first exposed to the association between pairs of nonsense words (e.g., "lif"/"neem") and pairs of objects. The word-object associations are introduced sequentially. After this exposure phase, participants perform a series of test trials. In each of these trials, one of the two sounds is uttered (e.g., “lif”) and participants choose the corresponding object from the two alternatives. An overview of the task is shown in Figure \@ref(fig:task).

From the general expression (1), we derive three exact analytical solutions instantiating different learning assumptions. The first solution is derived by assuming that the labels are recovered from sounds with a certain level of uncertainty $\sigma_L$, but that concepts are unambiguously recovered from the observed objects, i.e., $\sigma_C \rightarrow 0$. This assumption has been made --- whether implicitly or explicitly --- by most previous work in this line of research. One important implication of this assumption is that only the similarity of word sounds modulates success in word learning, not the similarity of the referents (as long as these referents are differentiated perceptually). This assumption yields the following probability function:

\begin{equation}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

The second solution is derived by making the more general assumption that both the labels and the concepts are recovered with ambiguity from the sounds and objects. We first introduce the simplifying assumption that the label-related uncertainty $\sigma_L$ and the concept-related uncertainty $\sigma_C$ are of a similar magnitude, i.e., $\sigma_C \approx \sigma_L = \sigma$. This assumption makes the prediction that the sound similarity and the object similarity impact word learning accuracy in exactly the same way. Furthermore, it allows us to study the behavior of the model with minimal free parameters.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

We finally derive the third (and most general) solution which allows label- and concept-related uncertainties to vary independently.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation} show some simulations of the accuracy $P(o_T|s)$ as a function of the distinctiveness parameters ($\Delta s$ and $\Delta o$) and the uncertainty parameters $\sigma_L$ and $\sigma_C$.

@@

The simulations explain two experimental results from previous studies and make one new prediction:

@@

2) For fixed values of $\Delta s$ and $\Delta o$, accuracy increases when the representational uncertainty $\sigma$ decreases. This fact may explain development, i.e., younger children have noisier representations [see @swingley2007; @yoshida2009], which leads to lower word recognition accuracy, especially for similar-sounding words.

3) For fixed values of $\Delta s$ and $\sigma$, accuracy increases with the visual distance between the semantic referents $\Delta o$. This is a new prediction that our model makes. Previous work studied the effect of several bottom-up and top-down properties in disambiguating similar sounding words [e.g., @fennell2010; @rost2009; @thiessen2007], but to our knowledge, no previous study in the literature tested the effect of the visual distance between the semantic referents.

@@

In this experiment, we tested participants in the word learning task introduced above (Figure \@ref(fig:task)). More precisely, we explored the predictions related to both distinctiveness and precision. Sound similarity ($\Delta s$) and object similarity ($\Delta o$) were varied simultaneously in a within-subject design. Two age groups (preschool children and adults) were tested on the same task to explore whether development can be characterized with the uncertainty parameters, $\sigma_C$ and $\sigma_L$. The experiment, sample size, exclusion criteria and the model's main predictions were pre-registered.\footnote{https://osf.io/942gv/}

@@

d_child <- read_delim("../Analysis/kids_all_final.txt", delim = " ") %>%
  filter(type == "Test") %>%
  filter(age != 6) %>%
  filter(code != 'test',
         !is.na(code)) %>%
  filter(grepl('soph', code)) %>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

@@

### Participants
We report data from $N=$ `r N_good_ch` children ages 4-5 years from the Bing Nursery School on Stanford University’s campus. An additional $N=$ `r N_all_ch-N_good_ch` children participated but were removed from analyses because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also report data from $N=$ `r N_good_ad` adult participants tested on Amazon Mechanical Turk. An additional $N=$ `r N_all_ad-N_good_ad` were tested but removed from analyses because they had low scores on the catch trials or because they were familiar with the non-English sound stimuli we used in the adult experiment.

### Stimuli and similarity rating
The sound stimuli were generated using the MBROLA Speech Synthesizer [@dutoit1996]. We generated three kinds of nonsense word pairs which varied in their degree of perceptual similarity to English speakers: 1) *different* pairs: "lif"/"neem" and "zem"/"doof", 2) *intermediate* pairs: "aka"/"ama" and "ada"/"aba", and 3) *similar* non-English pairs: "ada"/"a\textipa{d\super h}a" (in hindi) and "a\textipa{Q}a"/"a\textipa{\textcrh}a" (in arabic).

As for the objects, we used the Dynamic Stimuli javascript library\footnote{https://github.com/erindb/stimuli} which allowed us to generate objects in four different categories: "tree," "bird," "bug," and "fish." These categories are supposed to be naturally occurring kinds that might be seen on an alien planet. In each category, we generated *different*, *intermediate* and *similar* pairs by manipulating a continuous property controlling features of the category's shape (e.g, body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a separate survey on Amazon Mechanical Turk where we asked $N=20$ adults participants to evaluate the similarity of each sound and object pair on a 7-point scale. Data are shown in Figure \@ref(fig:stim) where we scaled responses within the range [0,1] for each stimulus group. These data will be used in all models as a proxy for the perceptual distance between the sound pairs ($\Delta s$) and the object pairs ($\Delta o$).

<!--Though these distances were collected from adults particiapants only, we make the assumption that they measure an perecptual distances represent the "objective" for both adults and children and that only the way these sounds are paired to object vary across age groups.-->

@@

#Import data from saved
data_human_model <- feather::read_feather("../saved_data/data_human_model.feather") %>%
  mutate(key = ifelse(key == "prediction_no", "Model 1",
                      ifelse(key == "prediction", "Model 2", "Model 3")))

@@

###Design

Each age group saw only two of the three levels of similarity described in the previous sub-section: *different* vs. *intermediate* for the preschoolers, and *intermediate* vs. *similar* for adults. We made this choice in light of pilot studies showing that adults were at ceiling with *different* sounds/objects, and children were at chance with the *similar* sounds/objects. That said, this difference in the level of similarity is accounted for in the model through using the appropriate distance (Figure \@ref(fig:stim)).

@@

To maximize our ability to measure subtle stimulus effects, the experiment was a 2x2 within-subjects factorial design with four conditions: high/low sound similarity crossed with high/low visual object similarity. Besides the four conditions, we also tested participants on a fifth catch condition which was similar in its structure to the other ones but was used only to select participants who were able to follow the instructions and show minimal learning.

@@

```{r allData, fig.env = "figure", fig.pos = "h", fig.align='center', fig.cap = "Accuracy of word recognition as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both the models' predictions (dashed lines) and the experimental results (solid lines). The single-variance model uses one joint fitting parameter for both sound and meaning variances. The double-variance model uses two separate fittings parameters for the sound and the meaning variances. Error bars represent 95\\% confidence intervals."}

@@

Experimental results are shown in Figure \@ref(fig:allData) (solid lines). We first analyzed the results using a mixed-effects logistic regression with sound distance, object distance and age group as fixed effects, and with a maximal random effects structure (allowing us to take into account the full nested structure of our data) [@barr2013]. We found main effects for all the fixed effects in the regression. For the sound distance, we obtained $\beta =$ `r s_coef` ($p$ `r p_val(s_p)`), replicating previous findings that sound distance modulates success in word learning [e.g., @stager1997]. For object distance, we found $\beta =$ `r c_coef`  ($p$ `r p_val(c_p)`), and this finding confirms the new prediction of our model according to which object distance also modulates success in word learning. Finally, for the age group, we obtained  $\beta =$ `r age_coef` ($p$ `r p_val(age_p)`), showing that overall performance improves with age.

@@

R2_model1 <-  round(cor(data_correlations$human, data_correlations$prediction_no)^2, 2)
R2_model2 <-  round(cor(data_correlations$human, data_correlations$prediction)^2, 2)
R2_model3 <-  round(cor(data_correlations$human, data_correlations$prediction_2)^2, 2)

mod0_val <- feather::read_feather("../saved_data/values_mod_no.feather") %>%
  rename(lower =`2.5 %`,
         upper = `97.5 %`)

@@

```{r models, results="asis", echo=FALSE}

mods <- data.frame(models = c('model 1', 'model 2', 'model 3'),
                   struc = c("Sig\\textsubscript{L} only",
                             "Sig\\textsubscript{L} = Sig\\textsubscript{C}",
                             "Sig\\textsubscript{L} != Sig\\textsubscript{C}"),
                   param = c(1,1,2),
                   R = c(R2_model1, R2_model2, R2_model3),
                   sigS_C = c(round(mod0_val$val[which(mod0_val$group=="children")],2),
                              round(mod1_val$val[which(mod1_val$group=="children")],2),
                              round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_s")],2)
                              ),
                   sigO_C = c('--',
                              round(mod1_val$val[which(mod1_val$group=="children")],2),
                              round(mod2_val$val[which(mod2_val$group=="children" & mod2_val$param=="Vr_o")],2)
                              ),
                   sigS_A = c(round(mod0_val$val[which(mod0_val$group=="adults")],2),
                              round(mod1_val$val[which(mod1_val$group=="adults")],2),
                              round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_s")],2)
                              ),
                   sigO_A = c('--',
                              round(mod1_val$val[which(mod1_val$group=="adults")],2),
                              round(mod2_val$val[which(mod2_val$group=="adults" & mod2_val$param=="Vr_o")],2)
                   ))

kable(mods, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Characteristics and performance of the models used in this study.",
      col.names = c("Model",  "Structure", "Param.", "R\\textsuperscript{2}", "Sig\\textsubscript{L}", "Sig\\textsubscript{C}", "Sig\\textsubscript{L}", "Sig\\textsubscript{C}")) %>%
  add_header_above(c("", "","","","Children" = 2, "Adults" = 2)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

We next fit the three models obtained through expressions (1), (2), and (3) to the participants' responses in each age group. The predictions of the models are shown \@ref(fig:allData) (dashed lines) and the parameter estimates (for $\sigma_L$ and $\sigma_C$) as well as models' goodness to fit (i.e., measured through $R^2$) are presented in Table \@ref(tab:models).

Model 1, which does not take into account ambiguity in recovering concepts from observed objects, explains only a small part of the variance. In contrast, Model 3, which does take into account this ambiguity, accounts for all the variance. Interestingly, Model 2 which has a single, shared uncertainty parameter for both auditory and visual modalities still explains almost all the variance in human data, suggesting that the explanatory power of the general Model 3 is largely due to its structure, rather than its degrees of freedom.

As predicted, the uncertainty parameters were larger for children than they were for adults (Table \@ref(tab:models)), showing that the probabilistic representations become more refined (that is, $\sigma$ becomes smaller) across development. Further, the parameter estimates of Model 3 show that this developmental effect is more important for the label-specific uncertainty that it is for the concept-specific uncertainty.
